---
title: "Package Flow"
author: "Danilo Sarti"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Package Flow}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  comment = "#>"
)
```


# Loading Packages 
```{r, message=FALSE}
library(InnoVaR)
library(tidyverse)
library(ggplot2)
library(readxl)
library(readr)
library(copula)
library(mice)
library(checkmate)
library(jmuOutlier)
library(ggpubr)
```

# Loading datasets containing the attributes of the modules to be simulated

```{r}
load("/Volumes/T7 Touch/InnoVaR/data/gen_att.rda")
head(gen_att)

load("/Volumes/T7 Touch/InnoVaR/data/soil_att_ex.rda")
head(soil_att_ex)


```


# Simulating soil module:

## declaring global information
```{r}
## declaring the number of variables in the module
n_var_soil=ncol(soil_att_ex)
## filtering the categorical ones.
soil_cat=soil_att_ex %>% select(which(sapply(soil_att_ex, class) == 'character'))
## filtering the numerical values.
soil_num=soil_att_ex %>% select(which(sapply(soil_att_ex, class) != 'character'))
```

## Informing correlations known a priori

```{r}
known_neg_dis=calculate_trig_dis(c(-0.8016201, -0.8022176, -0.7633010))
known_pos_dis=calculate_trig_dis(c(0.5,0.6,0.9))
```

## creating a data frame to store 
```{r}
n_var=ncol(soil_att_ex)
soil_coordinates=module_coords(n_var=ncol(soil_att_ex), 
                          which_neg=c("2_1","6_2","1_9","1_6","2_3"),
                          dis_neg_know_ass=known_neg_dis,
                          coo_known_neg_ass=c("2_1","6_2","1_9"),
                          dis_pos_know_ass=known_pos_dis,
                          coo_known_pos_ass=c("3_9","4_5","6_8"),
                          by=0.001
                         )

```
## creating symetric matrices of trigonometric distance or sins. 
```{r}
## containing the 
sym_tri_soil=sym_mtx(vector=soil_coordinates$coords$distance, n_var=n_var,var_names=colnames(soil_att_ex))

sym_sin_soil=sym_mtx(vector=soil_coordinates$coords$sin, n_var=n_var,var_names=colnames(soil_att_ex))
```

## Assuring correlation matrices 

```{r}
soil_corr=round(assure_corr(n_var=n_var,corr=sym_sin_soil),2)
```

## Setting a copula to simulation


```{r}
sim_data_cop=set_sim_copula(d=n_var,
                lower_tri_corr=soil_corr[lower.tri(soil_corr,diag=FALSE)],
            n_cont_var=ncol(soil_num),
            cont_var_par=list(list(lambda=40),list(df=30,ncp=3),
                  list(shape=40,scale=10),list(rate=30)) ,
            n_unique=length(unique(na.omit(soil_att_ex$siteid))),
           mar_cont_dists=c("pois","chisq","gamma","exp"),
           var_names=colnames(soil_att_ex)
            )
```


```{r}
###
cut_data_soil=cut_levels(categorical=soil_cat,
                    cat_inc_all_levels=c("siteid","yes_no_sample"),
levels_inc_all=list(siteid=c(1:16),yes_no_sample=c(1:2)),                   continuous_to_cut=sim_data_cop$simulated
                    )
head(cut_data_soil)
```

## Amputing 
```{r}
#amp_data_soil=ampute(cut_data_soil,mech="MAR",prop = 0.25)
#head(amp_data_soil$amp)
```


# Simulating genomic module


## declaring global information

We need to simulate the I genotypes expression of n markers. We assume that the genetic constitution of the genotypes does not differ across environments.

```{r}
#our problem is to simulate the expression of 100 traits for each genotype inside each J environments. 
known_neg_dis_gen=calculate_trig_dis(c(-0.8016201, -0.8022176, -0.7633010,0.80))
known_pos_dis_gen=calculate_trig_dis(c(0.5,0.6,0.9,0.1))
# we redefine the structure
n_genotypes=10
n_markers= ncol(gen_att)
gen_i_expr=vector("list", length = n_markers)

## we need to insert the levels of genotypes to be considered 
genotypes_id=sprintf("genotype_%d",seq(1:n_genotypes))
gen_att=add_to_df(gen_att,genotypes_id)
names(gen_att)[101]="genotype_id"


gen_i_coords=module_coords(n_var=ncol(gen_att),
                           which_neg = c("1_2","4_5","2_3","20_4","30_3"),
                           dis_neg_know_ass =known_neg_dis_gen,
                        coo_known_neg_ass=c("4_5","2_3","20_4","30_3"),
                        dis_pos_know_ass =known_pos_dis_gen,
                        coo_known_pos_ass = c("3_9","4_5","6_8","7_8"), 
                        by=0.0001)


## creating the symetric matrices of distances and corrs
sym_tri_gen_i=sym_mtx(vector=gen_i_coords$coords$distance,
                    n_var=ncol(gen_att),
                    var_names = colnames(gen_att)
                    )
sym_sin_gen_i=sym_mtx(vector=gen_i_coords$coords$sin,
                    n_var=ncol(gen_att),
                    var_names = colnames(gen_att)
                    )
gen_i_corr=round(assure_corr(n_var=ncol(gen_att),corr=sym_sin_gen_i),2)

sim_data_cop_gen_i=set_sim_copula(d=ncol(gen_att),
                            lower_tri_corr = gen_i_corr[lower.tri(gen_i_corr,diag=FALSE)],
                            n_cont=0,
                            cont_var_par = list(),
                           n_unique=length(unique(na.omit(gen_att$genotype))),
                           
                          mar_cont_dists=c(),
         
           var_names=colnames(gen_att))

cut_data_gen_i=cut_levels(categorical=gen_att,
                    cat_inc_all_levels=c("genotype_id"),
levels_inc_all=list(genotype_id=c(1:10)),                   continuous_to_cut=sim_data_cop_gen_i$simulated,corrs=gen_i_corr
                    )

## this data contains the expression of the markers
## for each of n_genotypes
head(cut_data_gen_i[1:3,c(1,4,101)])
## the data will repeat across n_sites environments.
n_sites=length(na.omit(soil_att_ex$siteid))
sites_id=sprintf("siteid_%d",seq(1:n_sites))
gens_site_id=rep(sites_id,each=n_genotypes)

abemus_genomics=do.call("rbind", replicate(n_sites, cut_data_gen_i, simplify = FALSE))
abemus_genomics$siteid=gens_site_id
head(abemus_genomics[1:3,c(1,10,101,102)])  
tail(abemus_genomics[1:3,c(1,10,101,102)])
```

# Merging simulated and genomic simulated data.

```{r}
#head(cut_data_soil)
#head(abemus_genomics)

final_soil_gen<-merge(abemus_genomics,cut_data_soil, by="siteid")

head(final_soil_gen[c(1:3),c(1,5,100,102)])
```

# Simulating phenomics

```{r}
library(dplyr)
dat <- final_soil_gen %>%
  dplyr::mutate_if(is.character, as.factor)

```

```{r}
target <- sim_target(
   X_gene = dat %>% dplyr::select(`trait_1`:`trait_100`),
   X_env = dat %>% dplyr::select(`DranaigeClass`:`yes_no_sample`),
   method = "lasso", pars = list(lambda = 0.2, sigma = 1,k=3),
   marginal_mean = 5.3, marginal_sd = 1.5
 )

 dat$target<-target$target

 qplot(genotype_id, target, data = dat,
       geom= "boxplot", fill = genotype_id)+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ggtitle("Simulated values of yield (t/ha) with Lasso")+ylab("yield (t/ha)")
 qplot(siteid, target, data = dat,
       geom= "boxplot", fill = siteid)+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ggtitle("Simulated values of yield (t/ha) with Lasso")+ylab("yield (t/ha)")
 #interaction.plot(dat$genotype, dat$environment, dat$yield, fixed = TRUE)
 dat %>%
   group_by(genotype_id,siteid) %>%
   summarise(mean_yield = mean(target)) ->tips2
 tips2 %>%
   ggplot() +
   aes(x = siteid, y = mean_yield, color = genotype_id) +
   geom_line(aes(group = genotype_id)) +
   geom_point()+ggtitle("Interactions with Lasso")+ylab("mean_yield (t/ha)")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

 

```




